---
title: "Project 5"
author: "Biraj GC"
date: "3/31/2020"
output: html_document
---
You are working as a data scientist for the government in the city of Washington, D.C. Currently, Washington, D.C has a bike sharing system. People could rent a bike from one location and return it to a different place. You are given a historical usage pattern with weather data contained in the Excel workbook bike.csv (please download from canvas). Preview the document. You are asked to forecast bike rental demand in the capital bike share program.

The data source is from Kaggle at https://www.kaggle.com/c/bike-sharing-demand 

This dataset contains the following data fields:

datetime - hourly date + timestamp

season - 1 = spring, 2 = summer, 3 = fall, 4 = winter

holiday - whether the day is considered a holiday workingday - whether the day is neither a weekend nor holiday

weather - 1: Clear, Few clouds, Partly cloudy, Partly cloudy 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog

temp - temperature in Celsius

atemp - "feels like" temperature in Celsius

humidity - relative humidity

windspeed - wind speed

casual - number of non-registered user rentals initiated

registered - number of registered user rentals initiated

count - number of total rentals


```{r}
library(tree)
library(randomForest)
```
```{r}
rm(list = ls())
```

Load the dataset bike.csv. Preview the document into memory. Then split the data into a training set containing 2/3 of the original data (test set containing remaining 1/3 of the original data).
```{r}
#load the csv file
bike <- read.csv("Bike.csv")
head(bike)
```
```{r}
#visualize the stucture
str(bike)
```
```{r}
#split the data into train set and test set

set.seed(1) #For reproducibility purposes

index <- sample(1:nrow(bike), 0.7*nrow(bike))

train <- bike[index, ]

test <- bike[-index, ]

```


1.Build a tree model using function tree().
  A.The response is count and the predictors are season, holiday, workingday, temp, atemp, humidity, windspeed, casual, and registered.
```{r}
#let's build a tree model
#medv (median house value) is our target variable

tree.bike <-  tree(count ~ season + holiday + workingday + temp + atemp + humidity + windspeed + casual + registered, data = train)

summary(tree.bike)
```

  B. Perform cross-validation to choose the best tree by calling cv.tree().
```{r warning=FALSE}
#CV of tree model
cv.bike <- cv.tree(tree.bike)
cv.bike
```
  c. Plot the model results of 2) and determine the best size of the optimal tree.
```{r}
#plot the devience and size to fnd the optimal tree
plot(cv.bike$size, cv.bike$dev, type = "b")
```
#We choose 5 as the optimal number of nodes.

  D. Prune the tree by calling prune.tree() function with the best size found in 3).
```{r}
#the best model with level 5
prune.bike = prune.tree(tree.bike, best =5)
```
  E. Plot the best tree model.
```{r}
#plot the pruned tree
plot(prune.bike)
text(prune.bike, pretty = 0)
```

  F. Compute the test error using the test data set.
```{r}
#working on test Hset
yhat_cv <-  predict(prune.bike, newdata = test)
```

```{r}
#compute the MSE
mean((yhat_cv-test$count)^2)
```
2.Build a random forest model using function randomForest()

  A. The response is count and the predictors are season, holiday, workingday, temp, atemp, humidity, windspeed, casual, and registered.
```{r  warning=FALSE}
#build a random forest model
rf.bike <-  randomForest(count ~season + holiday + workingday + temp + atemp + humidity + windspeed + casual + registered, data = train, importance = TRUE)
```

  B. Compute the test error using the test data set.
```{r}
#predictions on the test set
yhat.rf <-  predict(rf.bike, newdata = test)
```

```{r}
#Mean squared error
mean((yhat.rf - test$count)^2)
```

  C. Extract variable importance measure using importance() function.
```{r}
#importance  of variable
importance(rf.bike)
```

  D. Plot the variable importance using function varImpPlot(). Which are the top 2 important predictors in this model?
```{r}
#identify the importance of the variable
varImpPlot(rf.bike)
```
#registerd and casual are the top 2 predictors.
 

