---
title: "Project2_GC_Biraj"
author: "Biraj GC"
date: "2/9/2020"
output: html_document
---
```{r}
#reads in mtcars.csv into mtcars.
mtcars <- read.csv("mtcars.csv")
#check the data.
head(mtcars)
summary(mtcars)
```
#1) Use the lm() function to perform a simple linear regression with the response mpg and the predictor hp.

```{r}
#Use linear model function on data file for y/outcome as mpg and x/prdictor as hp  
lm_mod1 <- lm(mpg~hp,data=mtcars)
#summarize the model
summary(lm_mod1)
```
#2) Is there a relationship between the target mpg and predator hp?

#Answer = Yes, The relationship between mpg and hp is defined by the equation,  mpg = 29.297928 -(0.064548*hp) ie. For every unit increase in hp, mpg reduces by 0.064548.


#3)How strong is the relationship between the response and predictor?

#Answer = The three stars besides p-value suggests that the its very low, so the the relationship between response and predictor is very strong.


#4)Is the relationship between mpg and hp positive or negative?

#Answer = Negative relationship. This is because the slope in the model is negative, i.e downwards.


#5 What is the predicted mpg associated with a horsepower (hp) of 100? What’s the 95% confidence interval for the predicted mpg?

#answer = 29.297928-(-0.064548*100)=22.843128
```{r}
#use confint function to find 95% confidence interval. 
confint(lm_mod1,level=.95)
```
#For 95% confident interval the upper interval is mpg = 31.95 - 0.047 hp, mpg = 26.63 - 0.081 hp for the lower interval. 

#6 Plot the response and the predictor and add the regression line using abline().
```{r}
#plot for hp and mpg
plot(mtcars$hp,mtcars$mpg)+
#add abline using linear model
abline(lm_mod1, lwd=1, col = "red")
```

#7 Perform a multiple linear regression with mpg as the response and the predictors cyl, disp, hp, wt, vs, and gear. Print out the results using summary() function.

```{r}
#Use linear model function on data file for y/outcome as mpg and x/predictor as cyl,disp,hp,wt,vs,gear  
lm_mod2 <- lm(mpg~cyl+disp+hp+wt+vs+gear,data=mtcars)
#summarize the model.
summary(lm_mod2)
```
#8 Is there a relationship between the predictors and the response?
#Answer = Yes, the relationship is defined by the equation 33.29705 -0.74909*cyl + 0.01733*disp -0.03352*hp -3.85230*wt + 0.70816*vs + 1.14054*gear.

#9 Which predictors appears to have a statistically significant relationship to the response?

#Answer = hp and wt seem to have a significant relationship, The three stars in WT shows that it has more stronger significance than hp.

#10 Use * symbols to fit linear regression models with interaction effects between hp and wt. Does this interaction appear to be statistically significant?

```{r}
#Use linear model function on data file for y/outcome as mpg and x/predictor as hp and wt)  
lm_mod3 <- lm(mpg~hp*wt,data=mtcars)
summary(lm_mod3)
```
#Answer = It is significant since the relationship between hp and wt has three stars i.e the p value is very low..
