---
title: "Project 3"
author: "Biraj GC"
date: "2/18/2020"
output: html_document
---
```{r}
#load the library into memory
library(caret)
library(MASS)
library(leaps)
library (bestglm)
library(ISLR)
```


PART I.
You are a data scientist in a top dealership group in USA. Your boss, Mr. Lou, asked you to analyze the motor trend car data. You are given a dataset containing fuel consumption and 10 aspects of automobile design and performance for 32 automobiles the file mtcars.csv.

You are asked to perform the following tasks by writing a script in rmd. 

#1. Load the dataset mtcars.csv. Preview the document into memory and convert column am to a factor using factor() function.

```{r}
#load the data
mtcars <- read.csv("mtcars.csv")
#preview mtcars.
head(mtcars)
```
```{r}
#convert am into factor.
mtcars$am <- as.factor(mtcars$am)
str(mtcars$am)
```

#2. Split the data into training set and test set. The training set contains the first 35 observations, the test set containing the remaining observations.
```{r}
#split the data into training set.
mtcars.train <- mtcars[1:35,]
#split the data into test set. 
mtcars.test <- mtcars[-(1:35),] 
```


#3. Build a logistic regression model with the response is am and the predictors are mpg, cyl, hp, and wt using glm() function
```{r}
#logistic regression model using glm() function.
mod_mtcars <- glm(am~mpg+cyl+hp+wt,data = mtcars.train, family = binomial)
```
#4. Compute the test error on the test data set using a confusion matrix. Is it a good model based on test error?
```{r}
#use pedict function to test mtcars.test uing mod_pred
mtcars_pred <- predict(mod_mtcars,newdata = mtcars.test, type = "response")
#preview the prediction probabilities 
mtcars_pred
```
```{r}
#Set glm.prob as 0 for all test rows
glm.prob <- rep("0",nrow(mtcars.test))
#if prediction > 0.5 set as 1
glm.prob[mtcars_pred > 0.5] = "1"
#print glm.prob to visualize the data
glm.prob
```
```{r}
#visualize the data inform of table using confusion matrix.
table(glm.prob, mtcars.test$am)
```
5 out of 6 predictions were made correct.

```{r}
#test error = 1-(correct prediction - total)
test_error <- 1-((5)/(5+1+0+0))# 1-(5/6)
test_error
```
#ans: Based on the error, this is a good model, However since the test data set was very small the result is still inconclusive. 

Part II.
You are working as a data scientist for the city of Washington D.C. government. Currently, Washington D.C. has a bike sharing system. People can rent a bike from one location and return it to a different place. You are given a historical usage pattern with weather data contained in Excel workbook bike.csv. You are asked to forecast bike rental demand in the capital bike share program.

This data section contains the following data fields:

datetime - hourly date + timestamp

season - 1 = spring, 2 = summer, 3 = fall, 4 = winter

holiday - whether the day is considered a holiday workingday - whether the day is neither a weekend nor holiday

weather - 1: Clear, Few clouds, Partly cloudy, Partly cloudy 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog

temp - temperature in Celsius

atemp - "feels like" temperature in Celsius

humidity - relative humidity

windspeed - wind speed

casual - number of non-registered user rentals initiated

registered - number of registered user rentals initiated

count - number of total rentals

```{r}
#clear memory
rm (list = ls())
```

#1. Build a linear model to forecast number of total rentals (count) using potential predictors, season, holiday, workingday, weather, atemp, and registered.

```{r}
Bike <- read.csv("Bike.csv")
head(Bike)
```

```{r}
#load the model using lm()
Mod_Rental <- lm(count~season+holiday+workingday+weather+atemp+registered, data = Bike)
#summarize the model.
summary(Mod_Rental)
```


#2. Perform best subset selection using bestglm() function based on BIC. What’s the best model based on BIC?
```{r}
#The target variable is at the last column
Xy <-  Bike[, c("season", "holiday", "workingday", "weather", "atemp", "registered","count")]

#Convert it into a data.frame
Xy <- data.frame(Xy)

```

```{r}
#perform best subset selection based on BIC
lm.best <-  bestglm(Xy, IC = "BIC")

#print out the best model
lm.best
```
# Answer: According to Best subset selection and BIC, The Best model is count ~ season + holiday + workingday + weather + atemp + registered

#3. Compute the test error of the best model based on BIC using LOOCV.
```{r}
# define training control by specifying LOOCV
train_control <- trainControl(method="LOOCV")

```
```{r}
# train the model
model <- train(count~season+holiday+workingday+weather+atemp+registered, data=Bike, trControl=train_control, method="lm")

# summarize results
summary(model)

```
#Answer: The Residual standard error is 35.26 on 10879 degrees of freedom, and Adjusted R-squared is 0.9621

#4. Calculate the test error of the best model based on BIC using 10-fold CV.
```{r}
# define training control by specifying LOOCV
train_control_2 <- trainControl(method="CV", number=10)

```


```{r}
# train the model
model_2 <- train(count~season+holiday+workingday+weather+atemp+registered, data=Bike, trControl=train_control_2, method="lm")

# summarize results
summary(model_2)
```
#Answer: The Residual standard error is 35.26 on 10879 degrees of freedom, and Adjusted R-squared is 0.9621

#5. Perform best subset selection using bestglm() function based on CV. What’s the best model based on CV?
```{r}
#perform best subset seletion based on cross validation
lm.cv = bestglm(Xy, IC = "CV")

#print out the best model
lm.cv
```
# Answer: According to Best subset selection on CV, The Best model is count ~ season + holiday + workingday + weather + atemp + registered

#6. Perform the backward stepwise selection using stepAIC() function. What’s the best model?
```{r}
#run a linear model uing lm() function with no predictors 
null <- lm(count ~ 1, data = Bike)
#run a linear model uing lm() function with all predictors 
full <- lm(count ~ season+holiday+workingday+weather+atemp+registered, data = Bike)
```
```{r}
#run stepwise backward selection calling function stepAIC
lm.bwd <-  stepAIC(full, direction = "backward",
                   scope = list(lower = null, upper = full))

#print the best model
lm.bwd
```
# Answer: According to AIC & stepwise selection, The Best model is count ~ season + holiday + workingday + weather + atemp + registered

