---
title: "Project 4"
author: "Biraj GC"
date: "3/5/2020"
output: html_document
---
```{r}
#data clear
rm(list = ls())
```

You are a data scientist in Jedi Analytics (JA). JA is a public company traded in the Stock exchange. The CEO of JA wants to invest in the real estate properties in the Boston area. You are given a dataset containing housing values in the suburbs of Boston in the file Boston.csv (download this from canvas).

You are asked to perform the following tasks by writing a script in Rmd. 

```{r}
#read in all the Library
library(caret)
library(MASS)
library(boot)
library(gam)
```


1.Read the dataset in Boston.csv into R. Call the loaded data Boston. Make sure that you have the directory set to the correct location for the data.
```{r}
#load in the data
boston <-  read.csv("Boston.csv")
```

```{r}
#data preview using head
head(boston)
```
```{r}
#data preview using summary
summary(boston)
```
```{r}
#data preview using Str.
str(boston)
```


2.The response is nox and the predictor is dis. Use the poly() function to fit a cubic polynomial regression to predict nox using dis. Report the regression output.
```{r}
#build linear model using poly() function.
linMod <- glm(nox ~ poly(dis,3), data = boston)
#summarize the model.
summary(linMod)
```
#The three stars on all of the degrees of 'dis' indicate that the p-value is very low. Hence, All of the degree of dis is an important predictor.

3. Your assistant data scientist, Tom Johnson, is considering predicting nox using dis as a predictor. He proposes models from degree 5, degree 4, and degree 3, and degree 2 polynomial regression. Please perform cross-validation using caret package to select the optimal degree for the polynomial and justify your answer.

```{r}
#build linear model using poly() function for second third fourth and fifth degree.
#2 Degree
Lm2 = glm(nox ~ poly(dis,2), data = boston)
#3 Degree
Lm3 = glm(nox ~ poly(dis,3), data = boston)
#4 Degree
Lm4 = glm(nox ~ poly(dis,4), data = boston)
#5 Degree
Lm5 = glm(nox ~ poly(dis,5), data = boston)
```

```{r}
#perform 10-fold cross validation on all of the degrees
cv.error_10_2 <- cv.glm(boston, Lm2, K = 10)
cv.error_10_3 <- cv.glm(boston, Lm3, K = 10)
cv.error_10_4 <- cv.glm(boston, Lm4, K = 10)
cv.error_10_5 <- cv.glm(boston, Lm5, K = 10)

#extract the error data.
cv.error_10_2$delta[1]
cv.error_10_3$delta[1]
cv.error_10_4$delta[1]
cv.error_10_5$delta[1]
```
#since the third degree has lowest error aamong all, it is the optimal degree for the polynomial.

4.Leonard just took the DSCI 356. You recommend that he perform the following GAM analysis.
  1.Predict nox using a smoothing spline of degree 3 in dis and a smoothing spline of degree 2 in medv.
  2.Predict nox using a smoothing spline of degree 2 in dis and a smoothing spline of degree 1 in medv.
```{r warning = FALSE}
#build gam model for nox using smoothing splins of 2nd degree for medv and 3rd degree for dis
gam1 <- gam(nox ~ s(dis,3) + s(medv,2), data = boston)
#build gam model for nox using smoothing splins of 2nd degree for medv and 3rd degree for dis
gam2 <- gam(nox ~s(dis,2) + medv, data = boston)

```
 3.Perform anova analysis. Recommend the best model and justify your answer.
```{r}
#use anova function
anova(gam1,gam2)
```

#since the p vaue is really low and has tree star significans on the gam2 we reject the nul hypothesis, making gam2 the best model.
